---
title: "Assignment 4 - Applying meta-analytic priors"
author: "Riccardo Fusaroli"
output: html_document
---

``` {r Setup, include=FALSE}

# loading libraries
library(pacman)
pacman::p_load(tidyverse, metafor, brms, readr)

# loading data
d <- read_csv("Ass4_data.csv")
MA_d <- read_delim("Ass4_MetaAnalysisData.tsv", delim = "\t")

# setup randomization options
set.seed(123)

# Restore the models
MA_m <- readRDS(file = "models/MA_m.rds")
MA_m0 <- readRDS(file = "models/MA_m0.rds")
MA_m1 <- readRDS(file = "models/MA_m1.rds")
MA_prior <- readRDS(file = "models/MA_prior.rds")
pitch_skeptic_m1 <- readRDS(file = "models/pitch_skeptic_m1.rds")

```


## Assignment 4

In this assignment we do the following:
  - we run a Bayesian meta-analysis of pitch variability in ASD, based on previously published literature
  - we analyze pitch variability in ASD in two new studies using both a conservative and a meta-analytic prior
  - we assess the difference in model quality and estimates using the two priors.

The questions you need to answer are:
  - What are the consequences of using a meta-analytic prior?
  - Evaluate the models with conservative and meta-analytic priors.
  - Discuss the effects on estimates. Discuss the effects on model quality.
  - Discuss the role that meta-analytic priors should have in scientific practice.
  - Should we systematically use them?
  - Do they have drawbacks?
  - Should we use them to complement more conservative approaches?
  - How does the use of meta-analytic priors you suggest reflect the skeptical and cumulative nature of science?



### Step by step suggestions

Step 1: Perform a meta-analysis of pitch variability from previous studies of voice in ASD
  - the data is available as Ass4_MetaAnalysisData.tsv
  - You should calculate Effect size (cohen's d) and Standard Error (uncertainty in the Cohen's d) per each study, using escalc() from the metafor package (also check the livecoding intro)
  - N.B. we're only interested in getting a meta-analytic effect size for the meta-analytic prior (and not e.g. all the stuff on publication bias). See a brms tutorial here:
    https://vuorre.netlify.com/post/2016/09/29/meta-analysis-is-a-special-case-of-bayesian-multilevel-modeling/ The formula is EffectSize | se(StandardError) ~ 1 + (1 | Paper). Don't forget prior definition, model checking, etc.
    - Write down the results of the meta-analysis in terms of a prior for step 2.

```{r Part 1, echo=FALSE}

# ASD(TD)_N - how many ASD(TD) participant

# when talking about pitch there is no difference between Hz and log scale

# PitchMeanASDvsTD is if it significant higher or lower or NS (not significant)

# making sure all our valuables are possible to use - no one is characters by accident 

MA_d <- MA_d %>%
  mutate(
    PitchVariabilityASD_Mean = as.numeric(PitchVariabilityASD_Mean),
    PitchVariabilityTD_Mean = as.numeric(PitchVariabilityTD_Mean),
    PitchVariabilityASD_SD = as.numeric(PitchVariabilityASD_SD),
    PitchVariabilityTD_SD = as.numeric(PitchVariabilityTD_SD)
  )

# The scale is not the same, however not important since we look at it within studies   some are Hz some are SD

#Making sure our dataset only contains rows where the paper is not NA 

MA_d <- MA_d %>%
  subset(!is.na(Paper))


#Focusing on the variability in pitch

MA_d <- escalc(measure = "SMD",
               n1i = TD_N, #The number of TD in the study
               n2i = ASD_N, #ASD paticipants in the study 
               m1i = PitchVariabilityTD_Mean, #mean in PV for TD
               m2i = PitchVariabilityASD_Mean, #mean in PV for ASD
               sd1i = PitchVariabilityTD_SD, # SD in PV for TD
               sd2i = PitchVariabilityASD_SD, #SD in PV for ASD 
               data = MA_d,
               slab = Paper) # if we make a plot, what should be the name 

#adding new columns: yi is the effect size and vi is the varience of the effectsize

#calculating the SE of the effectsize
MA_d <- MA_d %>% 
  mutate(StandardError = sqrt(vi)) %>% 
  rename(EffectSize = yi)


summary(MA_d$EffectSize) #explained at 18 min in the video


#Running an analysis 

MA_m <- bf(EffectSize | se(StandardError) ~ 1 + (1 | Population))

get_prior(MA_m, data =MA_d, family= gaussian())

MA_prior <- c(
  prior(normal(0,1), class = Intercept),# Sceptical, mean of 0 and how sceptical are we with the mean being 0
  prior(normal(0, .3), class = sd) # saying the variablility up to .6 but we expect it to be more likely to be smaller then that 
  )

MA_m0 <- brm(
  MA_m,
  data = MA_d,
  family = gaussian(),
  prior = MA_prior,
  sample_prior = "only",
  chains = 2,
  cores = 2
)

pp_check(MA_m0, nsamples = 100) # acceptable

MA_m1 <- brm(
  MA_m,
  data = MA_d,
  family = gaussian(),
  prior = MA_prior,
  sample_prior = T,
  chains = 2,
  cores = 2
)

pp_check(MA_m1, nsamples = 100)

summary(MA_m1)


#Saving the values
MA_mean <- fixef(MA_m1)[[1]]
MA_se <- fixef(MA_m1)[[2]]

MA_hetrogeneity <- 0.32

# Save the models
saveRDS(MA_m, file = "models/MA_m.rds")
saveRDS(MA_m0, file = "models/MA_m0.rds")
saveRDS(MA_m1, file = "models/MA_m1.rds")
saveRDS(MA_prior, file = "models/MA_prior.rds")

```


Step 2: Analyse pitch variability in ASD in two new studies for which you have access to all the trials (not just study level estimates)
  - the data is available as Ass4_data.csv. Notice there are 2 studies (language us, and language dk), multiple trials per participant, and a few different ways to measure pitch variability (if in doubt, focus on pitch IQR, interquartile range of the log of fundamental frequency)
  - Also, let's standardize the data, so that they are compatible with our meta-analytic prior (Cohen's d is measured in SDs).
  - Is there any structure in the dataset that we should account for with random/varying effects? How would you implement that? Or, if you don't know how to do bayesian random/varying effects or don't want to bother, is there anything we would need to simplify in the dataset?

``` {r Part 2, echo=FALSE}

# subset to columns we need
test <- d %>% 
  select(ID, Language, Pitch_Mean, Pitch_Median, Pitch_SD, Pitch_IQR, Diagnosis)

# Standardize pitch IQR
d$Pitch_IQR_s <- scale(d$Pitch_IQR)

```


Step 3: Build a regression model predicting Pitch variability from Diagnosis.
  - how is the outcome distributed? (likelihood function). NB. given we are standardizing, and the meta-analysis is on that scale, gaussian is not a bad assumption. Lognormal would require us to convert the prior to that scale.
        Due to the fact that the data is scaled the data will be centered around 0
        
  - how are the parameters of the likelihood distribution distributed? Which predictors should they be conditioned on? Start simple, with Diagnosis only. Add other predictors only if you have the time and energy! no energy...
  
  - use a skeptical/conservative prior for the effects of diagnosis. Remember you'll need to motivate it.
  - Evaluate model quality. Describe and plot the estimates. 

``` {r Part 3, echo=FALSE}

pitch_m <- bf(Pitch_IQR_s ~ 0 + Diagnosis + (1|ID))
pitch_m1 <- bf(Pitch_IQR_s ~ 0 + Language + Language:Diagnosis + (1|ID))

get_prior(pitch_m, data = d, family = gaussian())

skeptic_prior <- c(
  prior(normal(0, 0.1), class = b),
  prior(normal(0,0.1), class =sd),
  prior(normal(.5,.3), class = sigma))

skeptic_prior_check <- brm(
  pitch_m,
  data = d,
  family = gaussian(),
  prior = skeptic_prior,
  sample_prior = "only",
  chains = 2,
  cores = 2
  )

pp_check(skeptic_prior_check, nsamples = 100)

pitch_skeptic_m <- brm(
  pitch_m,
  data = d,
  family = gaussian(),
  prior = skeptic_prior,
  sample_prior = T,
  chains = 2,
  cores = 2
  )

pp_check(pitch_skeptic_m, nsamples = 100)

#evaluate 
#check if we learn something
summary(pitch_skeptic_m)
plot(hypothesis(pitch_skeptic_m, "DiagnosisTD < 0"))
hypothesis(pitch_skeptic_m, "DiagnosisTD < 0")


## the complicated model

get_prior(pitch_m1, data = d, family =gaussian())

skeptic_prior_m1 <- c(
  prior(normal(0, .1), class = b, coef = "Languagedk"),
  prior(normal(0, .1), class = b, coef = "Languageus"),
  prior(normal(0, .1), class = b, coef = "Languagedk:DiagnosisTD"),
  prior(normal(0, .1), class = b, coef = "Languageus:DiagnosisTD"),
  prior(normal(0, .1), class = sd),
  prior(normal(0.5, .3), class = sigma)
)

skeptic_prior_check_m1 <- brm(
  pitch_m1,
  data = d,
  family = gaussian(),
  prior = skeptic_prior_m1,
  sample_prior = "only",
  chains = 2,
  cores = 2
  )

pp_check(skeptic_prior_check_m1, nsamples = 100)

pitch_skeptic_m1 <- brm(
  pitch_m1,
  data = d,
  family = gaussian(),
  prior = skeptic_prior_m1,
  sample_prior = T,
  chains = 2,
  cores = 2
  )

pp_check(pitch_skeptic_m1, nsamples = 100)


saveRDS(pitch_skeptic_m, file = "models/pitch_skeptic_m.rds")
saveRDS(pitch_skeptic_m1, file = "models/pitch_skeptic_m1.rds")


#Hypothesis testing
plot(hypothesis(pitch_skeptic_m1, "Languagedk:DiagnosisTD < 0"))
plot(hypothesis(pitch_skeptic_m1, "Languageus:DiagnosisTD < 0"))

#Comparing the models
loo_model_weights(pitch_skeptic_m, pitch_skeptic_m1)

```


Step 4: Now re-run the model with the meta-analytic prior
  - Evaluate model quality. Describe and plot the estimates. 


``` {r Part 4, echo=FALSE}
MA_mean <- fixef(MA_m1)[[1]]
#rounds off to -0.44

informed_prior <- c(
  prior(normal(-0.44, .1), class = b, coef = "Languagedk"),
  prior(normal(-0.44, .1), class = b, coef = "Languageus"),
  prior(normal(-0.44, .1), class = b, coef = "Languagedk:DiagnosisTD"),
  prior(normal(-0.44, .1), class = b, coef = "Languageus:DiagnosisTD"),
  prior(normal(0, .1), class = sd),
  prior(normal(0.5, .3), class = sigma)
)
informed_prior_check <- brm(
  pitch_m1,
  data = d,
  family = gaussian(),
  prior = informed_prior,
  sample_prior = "only",
  chains = 2,
  cores = 2
  )

pp_check(informed_prior_check, nsamples = 100)

pitch_informed_m <- brm(
  pitch_m1,
  data = d,
  family = gaussian(),
  prior = informed_prior,
  sample_prior = T,
  chains = 2,
  cores = 2
  )

pp_check(pitch_informed_m, nsamples = 100)
#evaluate 
#check if we learn something
summary(pitch_skeptic_m)

plot(hypothesis(pitch_informed_m, "Languagedk:DiagnosisTD < 0"))
plot(hypothesis(pitch_informed_m, "Languageus:DiagnosisTD < 0"))

saveRDS(informed_prior_check, file = "models/informed_prior_check.rds")
saveRDS(pitch_informed_m, file = "models/pitch_informed_m.rds")

```

Step 5: Compare the models
  - Plot priors and posteriors of the diagnosis effect in both models
  - Compare posteriors between the two models
  - Compare the two models (LOO)
  - Discuss how they compare and whether any of them is best.


``` {r Part 5, echo=FALSE}
plot(hypothesis(pitch_skeptic_m1, "Languagedk:DiagnosisTD < 0"))
plot(hypothesis(pitch_skeptic_m1, "Languageus:DiagnosisTD < 0"))

plot(hypothesis(pitch_skeptic_m1, "Languagedk:DiagnosisTD < 0"))
plot(hypothesis(pitch_skeptic_m1, "Languageus:DiagnosisTD < 0"))

loo_model_weights(pitch_skeptic_m, pitch_informed_m)
```

Step 6: Prepare a nice write up of the analysis and answer the questions at the top.









